Observing the United States' Homelessness Crisis through Mathematical Methods

### 2024-03-04 paper.tex
part 1 complete

@@

\noindent
\textit{The National GDP will have a strong-enough correlation with housing availability to aid in accurately
accounting for cyclical economic trends.} \\

\noindent
Justification: We fully recognize that national economic health will not be inherently causal with housing availability
in select cities, but on a macroeconomic scale, the strength of the economy will be reasonably provide a benchmark for
predicting the economic decisions of housing construction companies and consumers.
This relies on the assumption that in times of economic prosperity,

@@

\begin{table}[H]
  \centering
  \begin{tabular}{|c c c c|} 
    \hline 
    Year & Total housing units & Occupied units & Vacant units \\ [0.5ex]
    \hline
    2010  &  234,891 & 217,256 & 17,635 \\
    2011  &  237,735 & 220,060 & 17,675 \\
    2012  &  239,718 & 222,584 & 17,134 \\
    2013  &  240,277 & 222,491 & 17,786 \\
    2014  &  240,961 & 222,868 & 18,093 \\
    2015  &  241,326 & 222,098 & 19,228 \\
    2016  &  242,070 & 221,320 & 20,750 \\
    2017  &  243,402 & 221,119 & 22,283 \\
    2018  &  244,382 & 222,748 & 21,634 \\
    2019  &  245,476 & 224,166 & 21,310 \\
    2020  &  247,926 & 229,701 & 18,225 \\
    2021  &  252,924 & 236,191 & 16,733 \\
    2022  &  255,178 & 239,800 & 15,378 \\ [1ex] 
    \hline
  \end{tabular}
  \caption{Housing Statistics in Albequerque, New Mexico \cite{Census2010ACSDP1Y2010.DP04}}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{|c c c c|}
    \hline
    Year & Total housing units & Occupied units & Vacant units \\ [0.5ex]
    \hline
    2010 & 302,465 & 280,453 & 22,012 \\
    2011 & 304,164 & 282,480 & 21,684 \\
    2012 & 306,694 & 285,476 & 21,218 \\
    2013 & 309,205 & 288,439 & 20,766 \\
    2014 & 311,286 & 290,822 & 20,464 \\
    2015 & 315,950 & 296,633 & 19,317 \\
    2016 & 322,795 & 304,157 & 18,638 \\
    2017 & 334,739 & 314,850 & 19,889 \\
    2018 & 344,503 & 323,446 & 21,057 \\
    2019 & 354,475 & 331,836 & 22,639 \\
    2020 & 367,337 & 344,629 & 22,708 \\
    2021 & 362,809 & 337,361 & 25,448 \\
    2022 & 372,436 & 345,246 & 27,190 \\ [1ex]
    \hline
  \end{tabular}
  \caption{Housing Statistics in Seattle, Washington \cite{Census2010ACSDP5Y2010.DP04}}
\end{table}

\iffalse
\begin{table}[H]
  \centering
  \begin{tabular}{|c|l|c|l|}
    \hline
    Year & GDP & Year & GDP \\ [0.5ex]
    \hline
    1980 & 2857.3085 & 2005 & 13039.197 \\
    1981 & 3207.04125 & 2006 & 13815.583 \\
    1982 & 3343.78925 & 2007 & 14474.227 \\
    1983 & 3634.0365 &  2008 & 14769.86175 \\
    1984 & 4037.614 & 2009 & 14478.06725 \\
    1985 & 4338.9805 & 2010 & 15048.971 \\
    1986 & 4579.6325 & 2011 & 15599.7315 \\
    1987 & 4855.21625 & 2012 & 16253.97 \\
    1988 & 5236.438 & 2013 & 16880.68325 \\
    1989 & 5641.5795 & 2014 & 17608.13825 \\
    1990 & 5963.1445 & 2015 & 18295.019 \\
    1991 & 6158.12925 & 2016 & 18804.91325 \\
    1992 & 6520.32725 & 2017 & 19612.1025 \\
    1993 & 6858.5585 & 2018 & 20656.5155 \\
    1994 & 7287.2365 & 2019 & 21521.395 \\
    1995 & 7639.74925 & 2020 & 21322.9495 \\
    1996 & 8073.12175 & 2021 & 23594.03075 \\
    1997 & 8577.5525 & 2022 & 25744.10825 \\
    1998 & 9062.81675 & 2023 & 27357.842 \\
    1999 & 9631.17175 & \\ 
    2000 & 10250.952 & \\
    2001 & 10581.929 & \\
    2002 & 10929.10825 & \\
    2003 & 11456.4495 & \\
    2004 & 12217.19575 & \\ [1ex]
    \hline
  \end{tabular}
  \caption{Gross Domestic Product of the United States in Billions of Dollars, Adjusted for Inflation \cite{FREDGDP}}
\end{table}
\fi

@@

@@

@@

>>>>

\usepackage{graphicx}
\graphicspath{{./figures/}}

@@

@@

\begin{table}[H]
  \centering
  \begin{tabular}{|c c c c|} 
    \hline 
    Year & Total housing units & Occupied units & Vacant units \\ [0.5ex]
    \hline
    2010  &  234,891 & 217,256 & 17,635 \\
    2011  &  237,735 & 220,060 & 17,675 \\
    2012  &  239,718 & 222,584 & 17,134 \\
    2013  &  240,277 & 222,491 & 17,786 \\
    2014  &  240,961 & 222,868 & 18,093 \\
    2015  &  241,326 & 222,098 & 19,228 \\
    2016  &  242,070 & 221,320 & 20,750 \\
    2017  &  243,402 & 221,119 & 22,283 \\
    2018  &  244,382 & 222,748 & 21,634 \\
    2019  &  245,476 & 224,166 & 21,310 \\
    2020  &  247,926 & 229,701 & 18,225 \\
    2021  &  252,924 & 236,191 & 16,733 \\
    2022  &  255,178 & 239,800 & 15,378 \\ [1ex] 
    \hline
  \end{tabular}
  \caption{Housing Statistics in Albequerque, New Mexico \cite{Census2010ACSDP1Y2010.DP04}.}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{|c c c c|}
    \hline
    Year & Total housing units & Occupied units & Vacant units \\ [0.5ex]
    \hline
    2010 & 302,465 & 280,453 & 22,012 \\
    2011 & 304,164 & 282,480 & 21,684 \\
    2012 & 306,694 & 285,476 & 21,218 \\
    2013 & 309,205 & 288,439 & 20,766 \\
    2014 & 311,286 & 290,822 & 20,464 \\
    2015 & 315,950 & 296,633 & 19,317 \\
    2016 & 322,795 & 304,157 & 18,638 \\
    2017 & 334,739 & 314,850 & 19,889 \\
    2018 & 344,503 & 323,446 & 21,057 \\
    2019 & 354,475 & 331,836 & 22,639 \\
    2020 & 367,337 & 344,629 & 22,708 \\
    2021 & 362,809 & 337,361 & 25,448 \\
    2022 & 372,436 & 345,246 & 27,190 \\ [1ex]
    \hline
  \end{tabular}
  \caption{Housing Statistics in Seattle, Washington \cite{Census2010ACSDP5Y2010.DP04}.}
\end{table}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\textwidth]{chart1}
  \caption{The number of vacancies in Albequerque and Seattle from 2010 to 2022 \cite{Census2010ACSDP1Y2010.DP04,Census2010ACSDP5Y2010.DP04}.}
\end{figure}

\noindent
Let $H(x)$ equal the number of available houses for a given year, where $x$ is the current year (absolute from year 0, such as 2024). Simple regression on this dataset results in a total trendline of

$$H_{\textrm{\small TOTAL}}(x) = 385x - 735,364$$

\noindent
where $R^{2}_{\textrm{\small TOTAL}} = 0.586$, and an average trendline of

$$H_{\textrm{\small AVG}}(x) = 192x - 367,682$$

\noindent
where $R^{2}_{\textrm{\small AVG}} = 0.586$. \\

\noindent
A coefficient of determination of $0.586$ is undeniably subpar due to the immense indifficulty in representing vast differences across the trends with linear functions.

@@

With this model, we can predict the following housing statistics for 2034 (10 years in the future), 2044 (20 years in the future), and 2074 (50 years in the future):

\begin{table}[H]
  \centering
  \begin{tabular}{|c c|c c|}
    \hline
    Year & Years from 2024 & Total Available Homes & Average Available Homes in Each City \\
    \hline
    2034 & 10 & 47,726 & 22,846 \\
    2044 & 20 & 51,576 & 24,766 \\
    2074 & 50 & 63,126 & 30,526 \\
    \hline
  \end{tabular}
\end{table}

@@

Overall, our model proves generally weak at predicting immense, unforseen changes to the economic situation of both cities such as the advent of war, the outbreak of a major epidemic/pandemic, or a large-scale natural disaster. This model is best for serving as a general, aggregate predictor of housing availability based on cyclical economic trends. Our coefficient of determination $R^{2}$ is certainly low, indicating a lack of accuracy, but this is certainly due to the difficulty of modeling previously-mentioned large-scale events with a linear model. \\

\noindent
Altogether, we can conclude that this model's impact serves best as a general guideline for understanding realistic public policy because it is solely based largely on aggregate, cyclical economic factors that can be reasonably relied upon to occur continuously.
